Nondeterministic Bigraphical Reactive Systems for Markov Decision Processes\thanks{Supported by organization x.}

### 2018-08-08 paper.tex
Almost acceptable section on Agent 1, and the technical bits of Agent 2

\providecommand\longrightarrowRHD{\relbar\joinrel\relbar\joinrel\mathrel\RHD}
\providecommand\longrightarrowrhd{\relbar\joinrel\relbar\joinrel\mathrel\rhd}

@@

\begin{document}
\title{Nondeterministic Bigraphical Reactive Systems for Markov Decision
  Processes\thanks{Supported by organization x.}}
\titlerunning{Nondeterministic BRSs for MDPs}
\author{Paulius Dilkas \and Michele Sevegnani}
\authorrunning{P. Dilkas \& M. Sevegnani}
\institute{University of Glasgow, Glasgow, UK}
\maketitle

@@

\begin{figure}
  \includegraphics{models/example_bigraph/foo_lhs.pdf}
  \caption{An example bigraph.}
  \label{example_bigraph}
\end{figure}

Consider the graphical representation of a bigraph in Fig.
\ref{example_bigraph}. The two white dashed rectangles represent \emph{regions},
i.e. parents of any otherwise parentless nodes. Grey dashed rectangles are
called \emph{sites} and encode parts of the model that have been abstracted
away. Nodes, sites, and regions are also known as \emph{places}. Two places with
the same parent, or two points with the same link are called \emph{siblings}.

@@

\begin{figure}
  \begin{tabular}{p{0.5\textwidth}p{0.5\textwidth}}
    \begin{minipage}{0.5\textwidth}
      \vspace{13mm}
      \lstinputlisting[caption=BigraphER code., captionpos=b, language=Big, label=example_listing]{models/example.big}
    \end{minipage}
    &
      \begin{minipage}{0.5\textwidth}
        \centering
        \includegraphics{models/example_ts.pdf}
        \caption{The full transition system.}
        \label{example_ts}
      \end{minipage}
  \end{tabular}
\end{figure}

@@

% TODO: introduce this
\begin{figure}
  \includegraphics{images/interface.jpg}
  \caption{The BigraphER Jupyter interface with syntax highlighting.}
  \label{interface}
\end{figure}

@@

Akin to \cite{dblp:conf/nfm/giaquintahimn18}.

\subsection{\#1}

Grid layout (but could be any nodes-and-edges layout, as long as it's clear how
to go back home). Tracking visited/unvisited states. Each unvisited state has a
probability $p$ of containing an object. The agent returns (and stays) home
after collecting a set number of objects (easily customisable). Starting at the
bottom left cell.

@@

We begin by introducing controls: the types of nodes, what they contain, and
what they are linked to. The full list of controls is in Listing
\ref{agent1_controls}. \texttt{Cell} represents a discrete location. It is
linked to either \texttt{Visited} or \texttt{Unvisited} (initially, all except
one cells are unvisited). \texttt{Cell}s always contain
\texttt{Directions}, and one \texttt{Cell} contains an \texttt{Agent}. The
\texttt{Agent} may contain an \texttt{Object}, which may contain another
\texttt{Object}, etc. \texttt{Directions} always contain four nodes:
\texttt{North}, \texttt{East}, \texttt{West}, \texttt{South}. A \texttt{West}
node can be linked either to one of the two \texttt{Void}s or to an
\texttt{East} node of a neighbouring \texttt{Cell} (and similarly with other
directions). \texttt{Void}s represent the edges of the map. Using two
\texttt{Void} controls prevents us from unintentionally connecting the opposite
edges of the map. \texttt{Void1} corresponds to top and left edges, while
\texttt{Void2} is for bottom and right.

@@

Next, we define two predicates: \texttt{home} and \texttt{goal} (see Listing
\ref{agent1_code}). The former states that the \texttt{Agent} is home
if it is in the southwest corner of the grid,
while the latter says that the agent should go home as soon as
it gets one object. A larger number of objects can be set by,
e.g., \texttt{Agent.Object.Object}. Additionally, Listing \ref{agent1_code}
defines the initial $2 \times 2$ grid, as described previously.

\lstinputlisting[caption=The action of going north, captionpos=b,
language=Big, label=agent1_north, firstline=88,
lastline=117]{models/agent1.big}

Lastly, we have five actions (four for the directions, and one for
going home). Each of the actions for directions carries three reaction
rules: one for going to an already visited cell, one for visiting a
cell for the first time, and one for visiting a cell for the
first time and finding an object. We will use the rules for going north
in Listing \ref{agent1_north} as an example. When going to an already visited
cell, the \texttt{Agent} simply traverses a link between the \texttt{North} node
of its previous \texttt{Cell} and the corresponding \texttt{South} node. As this
is the only reaction rule for this situation, its probability is $1$. When
going to an unvisited node, we equip an \texttt{Object} with probability $p$.
Regardless of whether that happens or not, the \texttt{Cell} changes its
link from \texttt{Unvisited} to \texttt{Visited}.

\lstinputlisting[caption=Reaction rule to stay home if a required number of
objects is acquired., captionpos=b, language=Big, label=agent1_home,
firstline=180, lastline=183]{models/agent1.big}

The action \texttt{home} has five reaction rules: going west/south to a
visited/unvisited cell that are just like the rules in Listing
\ref{agent1_north}, except every mention of \texttt{Agent} is replaced with
\texttt{goal}, and a rule to stay home if the goal is satisfied (in Listing
\ref{agent1_home}).

>>>>

\providecommand\longrightarrowRHD{\relbar\joinrel\relbar\joinrel\mathrel\RHD}
\providecommand\longrightarrowrhd{\relbar\joinrel\relbar\joinrel\mathrel\rhd}
\providecommand\longrightdoublearrowRHD{\mathrel\LHD\joinrel\relbar\joinrel\relbar\joinrel\mathrel\RHD}

@@

\begin{document}
\title{Nondeterministic Bigraphical Reactive Systems for Markov Decision
  Processes\thanks{Supported by organization x.}}
\titlerunning{Nondeterministic BRSs for MDPs}
\author{Paulius Dilkas}
\authorrunning{P. Dilkas}
\institute{University of Glasgow, Glasgow, UK}
\maketitle

@@

\begin{figure}
  \centering
  \includegraphics{models/example_bigraph/foo_lhs.pdf}
  \caption{An example bigraph.}
  \label{example_bigraph}
\end{figure}

Consider the graphical representation of a bigraph in Fig.
\ref{example_bigraph}. The white ellipses are the \emph{nodes}. Each node has a
type, called \emph{control}, denoted here by the labels \textsf{A}, \textsf{B},
and \textsf{C}. The two white dashed rectangles represent \emph{regions},
i.e. parents of any otherwise parentless nodes. Grey dashed rectangles are
called \emph{sites} and encode parts of the model that have been abstracted
away. Nodes, sites, and regions are also known as \emph{places}. Two places with
the same parent, or two points with the same link are called \emph{siblings}.

@@

\begin{figure}
  \centering
  \begin{tabular}{p{0.5\textwidth}p{0.5\textwidth}}
    \begin{minipage}{0.5\textwidth}
      \centering
      \lstinputlisting[caption=BigraphER code., captionpos=b, language=Big, label=example_listing]{models/example.big}
    \end{minipage}
    &
      \begin{minipage}{0.5\textwidth}
        \centering
        \vspace{12mm}
        \includegraphics[width=\textwidth]{models/example_ts.pdf}
        \caption{The full transition system.}
        \label{example_ts}
      \end{minipage}
  \end{tabular}
\end{figure}

@@

% TODO: introduce this
\begin{figure}
  \centering
  \includegraphics{images/interface.jpg}
  \caption{The BigraphER Jupyter interface with syntax highlighting.}
  \label{interface}
\end{figure}

@@

Akin to \cite{dblp:conf/nfm/giaquintahimn18} and maybe \cite{koenig1998xavier}.

\subsection{Collecting Objects in a Grid}

% TODO: what is a link, mention that it's a hyperedge (at least one node)
Grid layout (but could be any nodes-and-edges layout, as long as it's clear how
to go back home). Tracking visited/unvisited states. Each unvisited state has a
probability $p$ of containing an object. The agent returns (and stays) home
after collecting a set number of objects (easily customisable). Starting at the
bottom left cell.

@@

We begin with a list of controls in Listing \ref{agent1_controls}. On the right
hand side of each equal sign is the number of ports that each node of that
control has, while the \texttt{\textbf{atomic}} keyword specifies the nodes that
cannot contain other nodes. \texttt{Cell}s represents discrete locations. They
are linked to either \texttt{Visited} or \texttt{Unvisited} (initially, all
except one cells are unvisited). \texttt{Cell}s always contain
\texttt{Directions}, and one \texttt{Cell} contains an \texttt{Agent}. The
\texttt{Agent} may contain an \texttt{Object}. In order to represent multiple
objects, it is convenient to put the second \texttt{Object} inside the first,
and so on. \texttt{Directions} contains a subset of the four nodes,
corresponding to the four possible directions of travel: \texttt{North},
\texttt{East}, \texttt{West}, \texttt{South}. We connect neighbouring
\texttt{Cell}s by linking the \texttt{West} node of one \texttt{Cell} to the
\texttt{East} node of another (or \texttt{North} to \texttt{South}).

@@

Next, we define two predicates: \texttt{home} and \texttt{goal} (see Listing
\ref{agent1_code}). The former states that the agent is home
if it is in the southwest corner of the grid,
while the latter specifies a goal of collecting one object. A larger number of
objects can be set by nesting \texttt{Object}s inside each other, e.g.,
\texttt{Agent.Object.Object}. Additionally, Listing \ref{agent1_code} defines
the initial $2 \times 2$ grid, as described previously.

\lstinputlisting[caption=The action of going north., captionpos=b,
language=Big, label=agent1_north, firstline=88,
lastline=117]{models/agent1.big}

Then, we have an action for each of the four directions, each with three
reaction rules: one for going to an already visited cell, one for visiting a
cell for the first time, and one for visiting a cell for the first time and
finding an object. We will use the rules for going north in Listing
\ref{agent1_north} as an example. When going to an already visited cell, the
\texttt{Agent} simply traverses a link between the \texttt{North} node of its
previous \texttt{Cell} and the corresponding \texttt{South} node. As this is the
only reaction rule for this situation, its probability is $1$. When going to an
unvisited node, we collect an \texttt{Object} with probability $p$, or make the
transition without getting an \texttt{Object} with probability $1-p$.
Regardless, the \texttt{Cell} changes its link from \texttt{Unvisited} to
\texttt{Visited}.

\lstinputlisting[caption=Reaction rule to stay home if a required number of
objects is acquired., captionpos=b, language=Big, label=agent1_home,
firstline=186, lastline=189]{models/agent1.big}

Lastly, we have three actions for after the agent acquires the specified number
of objects: two for going home west or south (since these are the two directions
necessary and sufficient to reach home), and one for staying home. The
former type of actions have two reaction rules each, depending on whether the
destination cell is visited or not. They are just like the rules in Listing
\ref{agent1_north}, except every mention of \texttt{Agent} is replaced with
\texttt{goal}. The rule to stay home is quite straightforward and is in
Listing \ref{agent1_home}.

\subsection{Layers of Abstraction}

\lstinputlisting[caption={The initial map: three consecutive cells, the midmost
of which has an inner room.}, captionpos=b, language=Big, label=agent2_initial,
firstline=15, lastline=19]{models/agent2/agent.big}

\lstinputlisting[caption={Two ways to generate a room, leaving one door either
open or closed.}, captionpos=b, language=Big, label=agent2_room,
firstline=21, lastline=48]{models/agent2/agent.big}

\begin{figure}
  \centering
  \begin{minipage}{0.45\textwidth}
    \centering
    \includegraphics[width=\textwidth]{models/agent2/goIn_lhs.pdf}
  \end{minipage}
  $\longrightdoublearrowRHD$
  \begin{minipage}{0.45\textwidth}
    \centering
    \includegraphics[width=\textwidth]{models/agent2/goIn_rhs.pdf}
  \end{minipage}
  \caption{Left to right: the reaction rule to go inside; right to left: the
    reaction rule to go outside.}
  \label{go_in_out}
\end{figure}
